---
title: "Social Sensitivity"
output: html_notebook
---

Estimating social sensitivity

```{r}
library(tidyverse)
library(rstan)
library(igraph)
library(cmdstanr)
library(posterior)
library(dplyr)
```



# Get data ready

```{r}
load("all_obs_unique_comb_temp.RData")
df.edges <- all_obs_unique_comb

#give ids a number
id_to_number <- data.frame(From = unique(c(df.edges$From,df.edges$To) ),To = unique(c(df.edges$From,df.edges$To) ), numb = as.numeric(as.factor(unique(c(df.edges$From,df.edges$To) ) ) ) )

df.edges<-left_join(df.edges, id_to_number %>% dplyr::select(From, numb), "From")
df.edges$from <- df.edges$numb
df.edges <- df.edges %>% dplyr::select(-numb)

df.edges<-left_join(df.edges , id_to_number %>% dplyr::select(To, numb), "To")
df.edges$to <- df.edges$numb
df.edges <- df.edges %>% dplyr::select(-numb)

df.edges <- df.edges %>% dplyr::select(-From) %>% dplyr::select(-To)

#add dyad id
df.edges$dyad <- paste0(df.edges$from,"_",df.edges$to)
df.edges$dyad <- as.numeric(as.factor(df.edges$dyad))

#add indicator if sender is a lower number compared to reciever 
df.edges$sr_indicator <- ifelse(df.edges$from < df.edges$to, 1,2)

#make sure there are no missing data
df.edges <- df.edges %>% tidyr::drop_na()

#take a look
df.edges

#save(df.edges, file = "df_edges.RData")
```

Fix dyad id 
-dyad is just the ordered pair of from and to actors
-sr_indicator is the ordering of the pair. this should allow us to get back to the from/to columns from dyad
```{r}
df.edges$dyad<- ifelse(df.edges$from < df.edges$to, paste0(df.edges$from,"_",df.edges$to) , paste0(df.edges$to,"_",df.edges$from) ) 
df.edges$dyad<- as.numeric(as.factor(df.edges$dyad))
df.edges$sr_indicator<- ifelse(df.edges$from < df.edges$to, 1 , 2 ) 

#check all is well
df.edges[df.edges$dyad==11,]
```

For scaling
```{r}
scale2sd <- function(x) {
  (x - mean(x, na.rm = TRUE)) / (2 * sd(x, na.rm = TRUE))
}
```



# AMEN model

### Model 1 - k2 scaled

Model with only main effects for the past social structures

```{r}

#prep the data
data_stan <-list(n_nodes = nrow(id_to_number),
                 n_dyads = max(df.edges$dyad),
                 dyad_id = df.edges$dyad,
                 scan_id = as.numeric(as.factor(df.edges$scanID)),
                 scan_N = length(unique(df.edges$scanID)),
                 sr_indicator = df.edges$sr_indicator,
                 N = nrow(df.edges),
                 sender_id = df.edges$from,
                 receiver_id = df.edges$to,
                 Y = df.edges$weight,
                 past_weight = as.numeric(scale2sd(df.edges$past_weight)),
                 past_aggro = as.numeric(scale2sd(df.edges$past_in_aggro)),
                 past_reci = as.numeric(scale2sd(df.edges$past_reciprical)),
                 past_trans = as.numeric(scale2sd(df.edges$past_transitive)),
                 past_trans_agg = as.numeric(scale2sd(df.edges$past_transitive_aggro)),
                 past_in_modu = as.numeric((df.edges$past_in_modu)),
                 past_overall_modu = as.numeric(scale2sd(df.edges$past_overall_modu)),
                 past_ndvi = as.numeric(scale2sd(df.edges$NDVI)),
                 send_receive = df.edges$sr_indicator,
                 troop = as.numeric(as.factor(df.edges$Troop)),
                 K=2,
                 troop_N = length(unique(df.edges$Troop)),
                 useVI = 0
                 )


mod <- cmdstan_model("AMEN_observed_data_reduce_wPin_predP.stan", cpp_options = list(stan_threads = TRUE))

parallel::detectCores()
Sys.setenv(STAN_NUM_THREADS = 80)  # or more, depending on your system

# Sample model briefly
amen_main <- mod$sample(
  data = data_stan,
  chains = 4,
  parallel_chains = 4,
  threads_per_chain = 19,
  iter_warmup = 1000,
  iter_sampling = 1000,
  init = 0.1,
  max_treedepth = 12,
  refresh = 10
)



amen_main$save_object(file = "fit_amen_obs_2015_predP.RDS")

# can be read back in using readRDS
#amen_main <- readRDS("fit_amen_obs_2015_predP.RDS")

test<-amen_main$draws(variables = c(
  "intercept", "b_weight", "b_reci", "b_aggro",
  "b_trans", "b_trans_agg", "b_in_modu"
))

library(bayesplot)
mcmc_trace(amen_main$draws(variables = c(
  "intercept", "b_weight", "b_reci", "b_aggro",
  "b_trans", "b_trans_agg", "b_in_modu"
)))

mcmc_trace(amen_main$draws(variables = c(
  "z_multi_effects[1,3]","z_multi_effects[2,3]","z_multi_effects[3,3]","z_multi_effects[4,3]"
)))

mcmc_trace(amen_main$draws(variables = c(
  "z_multi_effects[1,4]","z_multi_effects[2,4]","z_multi_effects[3,4]","z_multi_effects[4,4]"
)))

mcmc_trace(amen_main$draws(variables = c(
  "sigma_nodes","sigma_multi_effects"
)))


amen_main$summary(variables = c(
  "intercept", "b_weight", "b_reci", "b_aggro",
  "b_trans", "b_trans_agg", "b_in_modu"
))

```


```{r}
subset_samples <- amen_main$draws(format = "draws_df")

df_samples_main <- subset_samples %>% dplyr::select(contains("b_"))

library(reshape2)
colnames(df_samples_main) <- c("Gave grooming"  ,    "Received grooming"     ,    "Formed triadic closure (grooming)"     ,   "Was in same community" ,   "Received aggression"  , "Formed triadic closure (grooming & aggression)"  )
df_samples_main_melt<-melt(df_samples_main)

df_samples_main_melt$variable <- factor(df_samples_main_melt$variable, levels=c("Gave grooming"  ,    "Received grooming" ,     "Received aggression"    ,    "Formed triadic closure (grooming)"       , "Formed triadic closure (grooming & aggression)" ,   "Was in same community") )

library(ggridges)
ggplot(df_samples_main_melt %>% filter(variable != "b_overall_modu"), aes(y=variable, x=value) ) + geom_density_ridges(alpha=1, fill="skyblue", scale=0.95) + xlim(-0.1,0.45) + geom_vline(xintercept = 0, linetype="dashed", color="black") + theme_bw() + labs(y="Past social structure", x="Effect on current grooming") + theme(text = element_text(size = 20))

```

```{r}
subset_samples <- amen_main$draws(format = "draws_df")
model_p <- subset_samples %>% dplyr::select(contains("p_pred"))

residuals <- data_stan$Y - model_p  # Or logit(Y) - mu_hat for latent scale residuals

mean_residuals <- colMeans(residuals)   # posterior mean residuals
mean_residuals[1] #the acf plots of the two models are so similar, just checking that they are different
df_res <- data.frame(time= as.numeric(df.edges$date-min(df.edges$date)), resid = mean_residuals)
df_res <- df_res %>% arrange(time)

#ggplot(df_res, aes(x = time, y = resid)) +
#  geom_line() + geom_smooth(method = "loess") +
#  labs(title = "Residuals over Time")

acf(df_res$resid, lag.max = 20, main = "ACF of Residuals")
```

### Model 2 - k2 scaled ENV

The main model with adjustments for NDVI

```{r}

#prep the data
data_stan <-list(n_nodes = nrow(id_to_number),
                 n_dyads = max(df.edges$dyad),
                 scan_id = as.numeric(as.factor(df.edges$scanID)),
                 scan_N = length(unique(df.edges$scanID)),
                 N = nrow(df.edges),
                 sender_id = df.edges$from,
                 receiver_id = df.edges$to,
                 Y = df.edges$weight,
                 past_weight = as.numeric(scale2sd(df.edges$past_weight)),
                 past_aggro = as.numeric(scale2sd(df.edges$past_in_aggro)),
                 past_reci = as.numeric(scale2sd(df.edges$past_reciprical)),
                 past_trans = as.numeric(scale2sd(df.edges$past_transitive)),
                 past_trans_agg = as.numeric(scale2sd(df.edges$past_transitive_aggro)),
                 past_in_modu = as.numeric((df.edges$past_in_modu)), #as.numeric(scale2sd(df.edges$past_in_modu)),
                 past_overall_modu = as.numeric(scale2sd(df.edges$past_overall_modu)),
                 past_ndvi = as.numeric(scale2sd(df.edges$NDVI)),
                 dyad_id = df.edges$dyad,
                 send_receive = df.edges$sr_indicator,
                 troop = as.numeric(as.factor(df.edges$Troop)),
                 K=2,
                 troop_N = length(unique(df.edges$Troop)),
                 useVI=0
                 )


#fit small model
library(cmdstanr)
mod_ndvi <- cmdstan_model("AMEN_observed_data_NDIV_reduce_wPin_predP.stan", cpp_options = list(stan_threads = TRUE))

Sys.setenv(STAN_NUM_THREADS = 80)  # or more, depending on your system
amen_main_ndvi <- mod_ndvi$sample(data = data_stan,
                  chains = 4,
                  parallel_chains = 4,
                  threads_per_chain = 19, 
                  iter_warmup = 1000,
                  init = 0.1,
                  max_treedepth = 12,
                  iter_sampling = 1000,
                  refresh = 10)

amen_main_ndvi$save_object(file = "fit_amen_obs_NDVI_2015.RDS")

# can be read back in using readRDS
#amen_main_ndvi <- readRDS("fit_amen_obs_NDVI_2015.RDS")

library(bayesplot)
mcmc_trace(amen_main_ndvi$draws(variables = c(
  "intercept", "b_weight", "b_reci", "b_aggro",
  "b_trans", "b_trans_agg", "b_in_modu", "troopID","sigma_troop"
)))

mcmc_trace(amen_main_ndvi$draws(variables = c(
  "z_multi_effects[1,3]","z_multi_effects[2,3]","z_multi_effects[3,3]","z_multi_effects[4,3]"
)))

mcmc_trace(amen_main_ndvi$draws(variables = c(
  "z_multi_effects[1,4]","z_multi_effects[2,4]","z_multi_effects[3,4]","z_multi_effects[4,4]"
)))

mcmc_trace(amen_main_ndvi$draws(variables = c(
  "sigma_nodes","sigma_multi_effects"
)))


amen_main_ndvi$summary(variables = c(
  "intercept", "b_weight", "b_reci", "b_aggro",
  "b_trans", "b_trans_agg", "b_in_modu", "b_ndvi","b_ndvi_aggro","b_ndvi_weight",
   "b_ndvi_reci",
   "b_ndvi_trans" ,
   "b_ndvi_trans_agg" ,
   "b_ndvi_in_modu" , "troopID"
))



```

Estimated effects plots
```{r}


hist(as.numeric(scale(df.edges$NDVI)))
ci_ndvi<-HDInterval::hdi(as.numeric(scale(df.edges$NDVI)))
low_ndvi = ci_ndvi[1]
high_ndvi = ci_ndvi[2]


library(reshape2)

subset_samples_env<-as.data.frame(amen_main_ndvi$draws(variables = c(
  "intercept", "b_weight", "b_reci", "b_aggro",
  "b_trans", "b_trans_agg", "b_in_modu", "b_ndvi","b_ndvi_aggro","b_ndvi_weight",
   "b_ndvi_reci",
   "b_ndvi_trans" ,
   "b_ndvi_trans_agg" ,
   "b_ndvi_in_modu" 
),
    format = "draws_df")
)

df_samples_main_env_weight <- subset_samples_env %>% dplyr::select(b_weight, b_ndvi_weight)
df_samples_main_env_weight  <- df_samples_main_env_weight %>% mutate(b_weight_low_NDVI=(b_weight + b_ndvi_weight * low_ndvi),b_weight_high_NDVI=(b_weight + b_ndvi_weight * high_ndvi)) %>% dplyr::select(b_weight_low_NDVI, b_weight_high_NDVI)
df_samples_main_env_weight$parameter <- "Groom to"
df_samples_main_env_weight_melt <- melt(df_samples_main_env_weight, id.vars = c("parameter") )
df_char <- data.frame(variable = c("b_weight_low_NDVI","b_weight_high_NDVI" ), NDVI = c("low","high") ) 
df_samples_main_env_weight_melt<-left_join(df_samples_main_env_weight_melt,df_char, by="variable")


df_samples_main_env_reci <- subset_samples_env %>% dplyr::select(b_reci, b_ndvi_reci)
df_samples_main_env_reci  <- df_samples_main_env_reci %>% mutate(b_reci_low_NDVI=(b_reci + b_ndvi_reci * low_ndvi),b_reci_high_NDVI=(b_reci + b_ndvi_reci * high_ndvi)) %>% dplyr::select(b_reci_low_NDVI, b_reci_high_NDVI)
df_samples_main_env_reci$parameter <- "Groom from"
df_samples_main_env_reci_melt <- melt(df_samples_main_env_reci, id.vars = c("parameter") )
df_char <- data.frame(variable = c("b_reci_low_NDVI","b_reci_high_NDVI" ), NDVI = c("low","high") ) 
df_samples_main_env_reci_melt<-left_join(df_samples_main_env_reci_melt,df_char, by="variable")


df_samples_main_env_aggro <- subset_samples_env %>% dplyr::select(b_aggro, b_ndvi_aggro)
df_samples_main_env_aggro  <- df_samples_main_env_aggro %>% mutate(b_aggro_low_NDVI=(b_aggro + b_ndvi_aggro * low_ndvi),b_aggro_high_NDVI=(b_aggro + b_ndvi_aggro * high_ndvi)) %>% dplyr::select(b_aggro_low_NDVI, b_aggro_high_NDVI)
df_samples_main_env_aggro$parameter <- "Aggro from"
df_samples_main_env_aggro_melt <- melt(df_samples_main_env_aggro, id.vars = c("parameter") )
df_char <- data.frame(variable = c("b_aggro_low_NDVI","b_aggro_high_NDVI" ), NDVI = c("low","high") ) 
df_samples_main_env_aggro_melt<-left_join(df_samples_main_env_aggro_melt,df_char, by="variable")

df_samples_main_env_trans <- subset_samples_env %>% dplyr::select(b_trans, b_ndvi_trans)
df_samples_main_env_trans  <- df_samples_main_env_trans %>% mutate(b_trans_low_NDVI=(b_trans + b_ndvi_trans * low_ndvi),b_trans_high_NDVI=(b_trans + b_ndvi_trans * high_ndvi)) %>% dplyr::select(b_trans_low_NDVI, b_trans_high_NDVI)
df_samples_main_env_trans$parameter <- "Transitivity groom"
df_samples_main_env_trans_melt <- melt(df_samples_main_env_trans, id.vars = c("parameter") )
df_char <- data.frame(variable = c("b_trans_low_NDVI","b_trans_high_NDVI" ), NDVI = c("low","high") ) 
df_samples_main_env_trans_melt<-left_join(df_samples_main_env_trans_melt,df_char, by="variable")

df_samples_main_env_trans_agg <- subset_samples_env %>% dplyr::select(b_trans_agg, b_ndvi_trans_agg)
df_samples_main_env_trans_agg  <- df_samples_main_env_trans_agg %>% mutate(b_trans_agg_low_NDVI=(b_trans_agg + b_ndvi_trans_agg * low_ndvi),b_trans_agg_high_NDVI=(b_trans_agg + b_ndvi_trans_agg * high_ndvi)) %>% dplyr::select(b_trans_agg_low_NDVI, b_trans_agg_high_NDVI)
df_samples_main_env_trans_agg$parameter <- "Transitivity aggro"
df_samples_main_env_trans_agg_melt <- melt(df_samples_main_env_trans_agg, id.vars = c("parameter") )
df_char <- data.frame(variable = c("b_trans_agg_low_NDVI","b_trans_agg_high_NDVI" ), NDVI = c("low","high") ) 
df_samples_main_env_trans_agg_melt<-left_join(df_samples_main_env_trans_agg_melt,df_char, by="variable")

df_samples_main_env_in_modu <- subset_samples_env %>% dplyr::select(b_in_modu, b_ndvi_in_modu)
df_samples_main_env_in_modu  <- df_samples_main_env_in_modu %>% mutate(b_in_modu_low_NDVI=(b_in_modu + b_ndvi_in_modu * low_ndvi),b_in_modu_high_NDVI=(b_in_modu + b_ndvi_in_modu * high_ndvi)) %>% dplyr::select(b_in_modu_low_NDVI, b_in_modu_high_NDVI)
df_samples_main_env_in_modu$parameter <- "In community"
df_samples_main_env_in_modu_melt <- melt(df_samples_main_env_in_modu, id.vars = c("parameter") )
df_char <- data.frame(variable = c("b_in_modu_low_NDVI","b_in_modu_high_NDVI" ), NDVI = c("low","high") ) 
df_samples_main_env_in_modu_melt<-left_join(df_samples_main_env_in_modu_melt,df_char, by="variable")

#group them together
df_bind <-bind_rows(df_samples_main_env_weight_melt,df_samples_main_env_reci_melt,df_samples_main_env_aggro_melt,df_samples_main_env_trans_melt,df_samples_main_env_trans_agg_melt, df_samples_main_env_in_modu_melt)

library(ggridges)

# Define custom colors
custom_colors <- c("low" = "#66c2a5", "high" = "#238b45")  # Two shades of green

colnames(df_samples_main) <- c("Gave grooming"  ,    "Received grooming"     ,    "Formed triadic closure (grooming)"     ,   "Was in same community" ,   "Received aggression"  , "Formed triadic closure (grooming & aggression)"  )
df_samples_main_melt<-melt(df_samples_main)

# Manually set the order of the levels for the parameter factor
df_bind$parameter <- factor(df_bind$parameter, levels = c("Groom to" ,   "Groom from"  ,    "Aggro from" ,    "Transitivity groom" ,    "Transitivity aggro" ,"In community"  ))

df_bind$parameter2 <- NA
df_bind[df_bind$parameter=="Groom to",]$parameter2 <- "Gave grooming"
df_bind[df_bind$parameter=="Groom from",]$parameter2 <- "Received grooming"
df_bind[df_bind$parameter=="Aggro from",]$parameter2 <- "Received aggression"
df_bind[df_bind$parameter=="Transitivity groom",]$parameter2 <- "Formed triadic closure (grooming)"
df_bind[df_bind$parameter=="Transitivity aggro",]$parameter2 <- "Formed triadic closure (grooming & aggression)"
df_bind[df_bind$parameter=="In community",]$parameter2 <- "Was in same community"

df_bind$parameter2 <- factor(df_bind$parameter2, levels = c("Gave grooming" ,   "Received grooming"  ,    "Received aggression" ,    "Formed triadic closure (grooming)" ,    "Formed triadic closure (grooming & aggression)" ,"Was in same community"  ))

ggplot(df_bind, aes() ) + geom_density_ridges(aes(y=parameter2, x=value, fill=NDVI) , alpha = 0.7,scale=0.95) + xlim(-0.7,1.0) + geom_vline(xintercept = 0, linetype="dashed", color="black")+ 
  scale_fill_manual(values = custom_colors) +
  labs(y="Past social structure", x="Effect on current grooming")+ theme_bw()+ theme(text = element_text(size = 20))


```

```{r}
subset_samples_ndvi <- amen_main_ndvi$draws(format = "draws_df")
model_p <- subset_samples_ndvi %>% dplyr::select(contains("p_pred"))

residuals <- data_stan$Y - model_p  # Or logit(Y) - mu_hat for latent scale residuals

mean_residuals <- colMeans(residuals)   # posterior mean residuals
mean_residuals[1] #the acf plots of the two models are so similar, just checking that they are different
df_res <- data.frame(time= as.numeric(df.edges$date-min(df.edges$date)), resid = mean_residuals)
df_res <- df_res %>% arrange(time)

#ggplot(df_res, aes(x = time, y = resid)) +
#  geom_line() + geom_smooth(method = "loess") +
#  labs(title = "Residuals over Time")

acf(df_res$resid, lag.max = 20, main = "ACF of Residuals")
```


inter outputs

Check where nodes are in latent space
```{r}
#add data together
subset_samples_env_space<-as.data.frame(amen_main_ndvi$draws(variables = c(
  "mean_multi_effects"
),
    format = "draws_df")
)
df_post_space_1 <- subset_samples_env_space %>% dplyr::select(contains(",1]"))
df_post_space_2 <- subset_samples_env_space %>% dplyr::select(contains(",2]"))
df_post_space_3 <- subset_samples_env_space %>% dplyr::select(contains(",3]"))
df_post_space_4 <- subset_samples_env_space %>% dplyr::select(contains(",4]"))


#make a plot of the senders
x_sender <- df_post_space_1[1,]
y_sender <- df_post_space_2[1,]
x_receiver <- df_post_space_3[1,]
y_receiver <- df_post_space_4[1,]

df_latent_space_plot <- data.frame(x_sender=t(x_sender), y_sender=t(y_sender), x_receiver=t(x_receiver), y_receiver=t(y_receiver), numb = 1:72) #44

df_latent_space_plot <- left_join(df_latent_space_plot,id_to_number, by="numb")
colnames(df_latent_space_plot)[1:4] <- c("sender_x", "sender_y", "receiver_x","receiver_y")
#ggplot(df_latent_space_plot, aes(x=sender_x, y=sender_y, label=From)) + geom_text(color="blue") +
#  geom_text(aes(x=receiver_x, y=receiver_y), color="red" ) + theme_bw() +xlim(-0.5,0.5)+ylim(-0.5,0.5)

ggplot(df_latent_space_plot, aes(x=sender_x, y=sender_y, label=From)) + geom_text(color="blue") +
  geom_text(aes(x=receiver_x, y=receiver_y), color="red" ) + theme_bw() #+xlim(-0.5,0.5)+ylim(-0.5,0.5)


```


